---
title: "Benchmarking countries' COVID indicators"
author: "Imanuel Costigan"
date: Sys.Date()
output:
  md_document:
    variant: gfm
    toc: true
    
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, 
  dev = "ragg_png",
  fig.path = "fig/"
)
library(tidyverse)
```

# COVID benchmarking

This repository contains R code that I have used to benchmark a country against
other countries.I have benchmarked Australia against a selection of OECD 
countries.

## Setup

First let's set up some global variables:

```{r globals}
country_of_interest <- "AUS"
countries <- c(country_of_interest, "NZL", "USA", "GBR", "CAN", "DEU", "FRA", 
  "AUT", "BEL", "SGP", "DNK", "ISR", "ITA", "JPN", "KOR", "NLD", "FIN", "NOR", 
  "SWE", "ESP", "CHE")
countries <- countries[order(countries)]
smaller_set_of_countries <- c(country_of_interest, "CAN", "NZL", "SGP")
start_date <- as.Date("2021-09-01")
```

## Data sourcing 

All data is sourced from [Our World in Data](https://github.com/owid/covid-19-data/tree/master/public/data).

```{r source}
url <- "https://covid.ourworldindata.org/data/owid-covid-data.csv"
owid <- read_csv(url, show_col_types = FALSE)
```

## Data post-processing

Post-processing of this data set includes:

0. Translating the "wide" format OWID data set into a "long" format.
1. Filtering for countries that are of interest (in my case, a selection of OECD
   countries).
2. Filling missing values with the last available value.

```{r post_processing}
owid_long_tbl <-
  owid |>
  filter(
    iso_code %in% countries,
    date >= start_date
  ) |>
  select(
    iso_code,
    location,
    date,
    stringency_index,
    people_fully_vaccinated_per_hundred,
    total_boosters_per_hundred,
    new_tests_smoothed_per_thousand,
    new_cases_smoothed_per_million,
    # "Tests conducted per new confirmed case of COVID-19, given as a rolling
    #  7-day average (this is the inverse of positive_rate)"
    tests_per_case,
    hosp_patients_per_million, # stock, not flow.
    icu_patients_per_million,  # stock, not flow.
    new_deaths_smoothed_per_million
  ) |>
  fill(
    stringency_index,
    people_fully_vaccinated_per_hundred,
    total_boosters_per_hundred,
    new_tests_smoothed_per_thousand,
    new_cases_smoothed_per_million,
    tests_per_case,
    hosp_patients_per_million,
    icu_patients_per_million,
    new_deaths_smoothed_per_million
  ) |>
  pivot_longer(
    cols = !c("iso_code", "location", "date"),
    names_to = "indicator",
    values_to = "value"
  ) |>
  mutate(
    indicator = as_factor(indicator)
  )
```

The following compares the country of interest to a selection of the other 
countries since the `start_date` defined above.

```{r time-series, fig.width = 11, fig.height = 8}
owid_long_tbl |> 
  filter(iso_code %in% smaller_set_of_countries) |>
  ggplot(aes(x = date, y = value, colour = iso_code)) +
  geom_point(size = 0.75) +
  scale_colour_brewer(type = "qual", palette = "Set1") +
  facet_wrap(vars(indicator), scales = "free_y") +
  theme_light() +
  theme(legend.position = "bottom") +
  labs(
    title = "COVID stats",
    subtitle = "Time series of COVID stats for select countries",
    x = "Date",
    y = "Value",
    colour = "Country ISO code",
    caption = paste("Data sourced from Our World in Data.", 
      "Prepared by @imanuelcostigan.")
  )
```

And finally, we compare the country of interest to the other countries. We do
this on the penultimate date that is published by OWID to ensure data is 
available across all countries.

```{r boxplot, fig.width = 10, fig.height = 10}
benchmark_date <- (owid_long_tbl |> pull(date) |> max(na.rm = TRUE)) - 1
owid_last_snap <-
  owid_long_tbl |>
  filter(date == benchmark_date)
owid_country_of_interest <- 
  owid_last_snap |>
  filter(iso_code == country_of_interest)
owid_last_snap |>
  ggplot(aes(x = indicator, y = value)) + geom_boxplot()  +
  geom_boxplot(data = owid_country_of_interest, colour = "red") +
  facet_wrap(vars(indicator), scales = "free") +
  theme_bw() +
  theme(axis.text.x = element_blank()) +
  labs(
    title = "COVID benchmarking",
    subtitle = paste0("Comparing ", country_of_interest, " (red) to ",
      paste0(countries[countries != country_of_interest], collapse = ", ")),
    caption = paste0("Data sourced from Our World in Data. Benchmarked for ",
      benchmark_date, ". Prepared by @imanuelcostigan.")
  )

```

